
\section{Abstract}
The Fumi package is comprised of several commands for the purpose of text mining for Japanese text. 

Data mining is mostly deal with quantitative information in structured format. 
In recent years, advancements in computer performance and data analysis technology have increased the capacity to handle unstructured information. This type of data contains numbers, time and facts. An example of unstructured information in essays written by humans (natural language). 

This package contains morphological analyser JUMAN for Japanese, and case structure analyser KNP. The algorithms are developed by Kyoto University Department of Intelligence Science and Technology, Kurohashi and Kawahara lab. 
Since the results generated are saved in CSV file format, various analytical processes with mcommand can be carried out to create analytical models. 
Please refer to the official website of JUMAN and KNP for more details. 

\begin{itemize}
\item Morphological Analyzer System  JUMAN \url{http://nlp.ist.i.kyoto-u.ac.jp/EN/index.php?JUMAN}
\item Case Structure Analyzer System KNP \url{http://nlp.ist.i.kyoto-u.ac.jp/EN/index.php?KNP}
\end{itemize}

 

%本「View（眺）」パッケージは、複数のデータ視覚化用コマンドから構成されている。
%視覚化とは、データをグラフやチャートとして描画することをいい、
%データを概観したり、資料に掲載したりする際には重要な役割を果たす。

%本パッケージには、
%円グラフを描画する\verb|mpie.rb|コマンド、
%Sankeyダイアグラム（流量グラフ）を描画する\verb|msankey.rb|コマンド、
%グラフデータを汎用的な形式に変換する\verb|mgv.rb|コマンド
%が含まれている。

%CSV形式のデータを入力として与え(Ruby拡張ライブラリを用いている)、
%D3ライブラリ(Data-Driven Documents)で作成された視覚化アプリケーションsankey diagram
%(\url{http://bost.ocks.org/mike/sankey/})を利用した、Sankeyダイアグラムや、
%円グラフを組み込んだ単体のhtmlを出力する。

