

%\documentclass[a4paper]{jarticle}


\section{mknp.rb Parsing case structure using KNP \label{sect:mknp}}
\index{mknp@mknp}

Given a document in text file, KNP parses the case structure and saves the output in XML format. 
This command can process multiple text files, and can be executed in multi-processing mode based on compatibility of OS. 


\subsection{Format}
\begin{verbatim}
mknp.rb I= O= [P=] [mp=] [log=] [-mcmdenv] [--help]
\end{verbatim}

\begin{table}[htbp]
{\small
\begin{tabular}{ll}
\verb|I=|       & : Path name of the text document input. \\
\verb|O=|       & : Path name of the analysis results saved as XML file.   \\
\verb|P=|       & : Path name to save the direct output of KNP (default do not return output).  \\
\verb|mp=|      & : Number of parallel processes.  Default is 2.   \\
\verb|log=|     & :  Log file of processing error from KNP.  \\
\verb|-mcmdenv| & : Display MCMD message containing environment variables.  \\
                & :  Default returns warning and error message  (KG\_VerboseLevel=2). \\  
\verb|--help|    & : Display help  \\
\end{tabular} 
}
\end{table} 


\subsubsection*{Example of input file} 

The structure of text is preferred to include one sentence per line. 
The character code must be in UTF8 format. 

\begin{Verbatim}[baselinestretch=0.7,frame=single]
子どもはリンゴがすきです。
望遠鏡で泳ぐ少女を見た。
\end{Verbatim}


\subsubsection*{Example of output file} 

Results of case structure analysis is stored in XML file format. 

\begin{Verbatim}[baselinestretch=0.7,frame=single]
<?xml version='1.0' encoding='UTF-8'?>
<article id='test.txt'>
 <sentence id='0' text='子どもはリンゴがすきです。'>
  <chunk id='0' link='2' phraseType='格助詞句' caseType='ガ２格' phrase='子供' phraseTok='子
   <token id='0' class1='名詞' class2='普通名詞' word='子ども' orgWord='子ども' daiWord='子供
   <token id='1' class1='助詞' class2='副助詞' word='は' orgWord='は'/>
  </chunk>
  <chunk id='1' link='2' phraseType='格助詞句' caseType='ガ格' phrase='林檎' phraseTok='リン
   <token id='2' class1='名詞' class2='普通名詞' word='リンゴ' orgWord='リンゴ' daiWord='林檎
   <token id='3' class1='助詞' class2='格助詞' word='が' orgWord='が'/>
  </chunk>
  <chunk id='2' link='-1' phraseType='用言句' phraseTok='すきだ' rawPhrase='すきです。' phrase
   <token id='4' class1='形容詞' class3='ナ形容詞' class4='デス列基本形' word='すきだ' orgWord
   <token id='5' class1='特殊' class2='句点' word='。' orgWord='。'/>
  </chunk>
 </sentence>
:
\end{Verbatim}

The \hyperref[sect:mcaseframe]{mcaseframe.rb} command can be used to extract dependency relationship from this XML.  


\subsection{Examples}
\input{examples/tex/mknp_ex.tex}

%\end{document}

