

%\documentclass[a4paper]{jarticle}


\section{mnewdic.rb Output adjacent word pair candidate from corpus \label{sect:mnewdic}}
\index{mnewdic@mnewdic}

From text corpus  (large set of text files),  returns the candidate information of adjacent word pair to be registered in the dictionary. 

JUMAN contains a standard dictionary with  common words, for text mining in specific fields, interpretation of 1 phrase  may be divided into multiple words. 
    
\verb|mnewdic.rb| command analyze the given corpus based on the frequency of phrases with words pairs that appears at the same time, and output the list in CSV file. 

\verb|mjumandic.rb| command simply dictionary registration in the process of text mining.    


\subsection{Format}
\begin{verbatim}
mnewdic.rb [i=] [O=] [S=] [n=] [seed=] [-dai] [-mcmdenv] [--help]
\end{verbatim}

\begin{table}[htbp]
{\small
\begin{tabular}{ll}
\verb|i=|       & : Corpus file name  \\
\verb|O=|       & : Output directory name  \\
\verb|S=|       & : Minimum value of word pair appearance \\
\verb|n=|       & : Output sentence number for word pair  \\
\verb|seed=|    & : Seed of random number  \\
\verb|-dai|     & : Use a symbol to represent headwords  \\
\verb|-mcmdenv| & : Display MCMD message containing environment variables.  \\
                &    Default returns warning and error message  (KG\_VerboseLevel=2). \\
\verb|--help|   & : Display help   \\
\end{tabular} 
}
\end{table} 

\subsubsection*{Example of input file }

Given the corpus file specified at \verb|i=| parameter, one row corresponds to one sentence in text file. 


\begin{Verbatim}[baselinestretch=0.7,frame=single]
3年ぶりにウォークマンを買ったけど、育休中はあまり活躍の余地がないですね。
待機児童解消の方がいい気がするけど。
:
\end{Verbatim}

\subsubsection*{Example of Output File  }

In the directory specified at \verb|O=| parameter, 
the results are saved in \verb|words.csv| file and  \verb|corpus.csv| file  （when \verb|nkf| command is installed,  the character code converted from Shift JIS is also saved in the output in both files. 
 
\begin{Verbatim}[baselinestretch=0.7,frame=single]
見出し語,品詞,読み,カテゴリ,ドメイン,pid,word1,word2,freq
職場復帰,,,,,0,職場,復帰,31
授業参観,,,,,1,授業,参観,28
会議参加,,,,,2,会議,参加,26
:
\end{Verbatim}

\newpage

The description of each item in \verb|words.csv| file is shown below.  

\begin{table}[htbp]
{\small
\begin{tabular}{ll}
\verb|見出し語| & : 見出し語 \\  
\verb|品詞|     & : 品詞 \\  
\verb|読み|     & : 読み \\  
\verb|カテゴリ| & : カテゴリ \\  
\verb|ドメイン| & : ドメイン \\  
\verb|pid|      & : pid \\  
\verb|word1|    & : 語1 \\  
\verb|word2|    & : 語2 \\  
\verb|freq|     & : 出現頻度 \\  
\end{tabular} 
}
\end{table} 

Users can refer to \verb|corpus.csv| to check whether the word of registered candidates appeared in the text.  


\begin{Verbatim}[baselinestretch=0.7,frame=single]
pid,id,text
0,52,"神戸で始めての、育休後職場復帰セミナーを開催しました。"
0,317,"僕の知り合いは2人子どもを産んで、立て続けに産休＋育休を取って、職場復帰した。"
:
\end{Verbatim}




\subsection{Examples}
\input{examples/tex/mnewdic_ex.tex}

%\end{document}

