
%\begin{document}

\section{Specify Fields\label{sect:fieldname}}
MCMD reads the field names in the first row of CSV data; the field can also be specified with a field number without the field names. There are four options for handling the row of field names, which are: -nfn,-nfno,-nfni and -x. Its usage will be illustrated with examples as follows. In addition, note that the field number starts at 0 from the left such as 0, 1, 2.

\input{examples/tex/field_name1_ex.tex}

\subsection{Valid fieldnames}
Fieldnames can contain the characters stated as follows:

\begin{itemize}
\item Alphabetic characters (a-z, A-Z)
\item Numerals (0-9)
\item Multibyte characters (such as UTF-8)
\item Symbols
\end{itemize}

However, it is recommended to avoid using the following symbols. Using the symbols will not return an error; however, the symbols may be used as special characters in MCMD and the special character function may not be available if it is used as a field name.

\begin{itemize}
\item \verb|,| Comma
\item \verb|:| Colon
\item \verb|%| Percent
\item \verb|*| Asterisk
\item \verb|?| Question mark
\item \verb|&| And
\item \verb|\| Backslash
\item \verb|]| Square brackets, right
\item \verb|[| Square brackets, left
\end{itemize}

\subsection{Valid item number}
When specifying field number, the field numbers can be listed with a comma delimiter. Alternatively, it is also possible to specify the number at the end of the field name (add \verb|"L"|) or specify the range (\verb|-|).

For example, the argument \verb|0L| specifies the last field, and \verb|2L| specifies the 2nd field from the end (note that field number starts from 0). Furthermore, when \verb|0-5| is specified, six fields starting from 0 to 5 are selected, which is equivalent to \verb|0,1,2,3,4,5|.

\input{examples/tex/field_number_ex.tex}

\subsection{Input and output fields}
The \verb|f=| parameter is used to specify the field(s) in many commands. The format of \verb|f=| is defined as “ input field:output field”. If an output fieldname is not specified, the input fieldname will be used as the output fieldname. In addition, you can specify the \verb|-x| option to combine a fieldname with a number, like \verb|f=0:Quantity|.

\input{examples/tex/field_io_ex.tex}

\subsection{Wildcard}
The wildcard characters  \verb|"*"| and  \verb|"?"|  can be used to specify multiple field names.
The asterisk sign \verb|"*"| matches 0 or more characters, and the question mark \verb|"?"| matches a single character.
Note that the order of evaluation of wildcard characters follows the order of the fields in the input data.
For example, if the order of the fields in input data is \verb|A5,A3,A4,A2,A1|, the parameter \verb|f=A*| is evaluated as \verb|f=A5,A3,A4,A2,A1|.

\input{examples/tex/field_wc_ex.tex}

\subsection{Replace the name of an output field}
The special character \verb|"&"| specified in the output field name can be replaced with the current field name.
For example, the parameter \verb|f=abc:xx&xx| returns \verb|xxabcxx| as the output field name.
The \verb|"&"| character can be specified at any position as many times as required in the output field name.
However, the ampersand is a special character in shell which is interpreted as "background execution". Thus, it is necessary to escape and enclose the field name in double quotes when including \verb|"&"| in field name.

\input{examples/tex/field_rep_ex.tex}

\subsection{Notes on summation commands}
A summation command sums records in specified fields for each key field. When one row is output per key field, records to be output are indefinite for non-specified fields. Take the msum command, for example, where there are four fields: customer, date, product, and amount. To sum the amounts for each customer, you would specify \verb|msum k=customer f=amount|. Regarding the records in the date and product fields, the output is indefinite.
%\end{document}
