%\begin{document}

%\section{Format of MCMD\label{sect:format}}
The format for all commands is explained in this section. 
The format of all command references is shown as in the following example.

\begin{table}[htbp]
\begin{center}
\begin{tabular}{l}
\hline
{\large Format} \\
\verb/  mjoin k= [f=] [K=] [-n] [-N] m=|/ 
\hyperref[sect:option_i]{i=}
\hyperref[sect:option_o]{[o=]}
\hyperref[sect:option_nfn]{[-nfn]} 
\hyperref[sect:option_nfno]{[-nfno]}  
\hyperref[sect:option_x]{[-x]}
\verb|[--help]|
\verb|[--version]|\\

\\
{\large Parameters} \\
\verb|  k=   List of field name(s) to match with the input data |\\
\verb|      [join key break processing: character string in ascending order]. |\\
\verb|      The field(s) in the input data is specified at the K= parameter.|\\
\verb|      Field in the reference data will be combined with same field in the record. |\\
\verb|      If the field(s) specified at K= from reference file do not match any values,|\\
\verb|      it is treated as a NULL value.|\\
\verb|  f=  Specify the list of field name(s) to join from the reference file. |\\
\verb|      When the f= parameter is not defined, all fields except the key field(s) | \\
\verb|      in the reference file will be joined to the input file. |\\
\verb|      :   :|\\
\hline
\end{tabular} 
\end{center}
\end{table} 

Command options and parameter can be specified after the command name. The parameter takes a value and the parameter name and value is separated by an “=” sign. An option begins with one dash “-” or two dash “--”. Command options and parameters that are common in most commands include \verb|i=,-nfn|. The link to that section is appended in a separate section. The description of each parameter is described in the format below.

Parameters enclosed in square brackets like \verb|[f=]| means that it is optional. On the other hand, parameters that are not enclosed in square brackets like \verb|k=| means that it is a required parameter. Parameters separated by a vertical bar enclosed in square parentheses, such as \verb/[to=|size=]/ means that the command will only read either to= or size=  (e.g. \hyperref[sect:mbest]{mbest command}). On the other hand, if the parameters separated by \verb/"|"/ is not enclosed in square brackets such as \verb/m=|i=/, one of these parameters is required and must be specified (e.g. mjoin command. See above).

Further, parameters or options are only required when specifying certain options, there are also more complex parameter conditions which is described in the description field of each parameter.

%\end{document}
